---
title: "Replace it"
author: " Carlos van der Ley, Christine Kwon, Connor, Leonardo Gonzalez"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## Title (replace it)



## Authors

Carlos van der Ley, Christine Kwon, Connor, Leonardo Gonzalez

## Introduction



## Questions

**1. **

**2. **

**3. **

**4. **

## Data

### Data Introduction



### Load packages
```{r load-packages, message=FALSE}
library(tidyverse)
library(tidymodels)
library(forcats)
library(olsrr)
library(skimr)
library(srvyr)
library(osfr) #from osf website
```

### Data Import
```{r}

# data <- list.files(recursive = TRUE,
#                         path = "testdata",
#                         pattern = "*.csv",
#                         full.names = TRUE) |>
#   purrr::map(~read_csv(.))
# 
# 
# save(data, file="data/mental_health.rda")

```

### Data Wrangling

#### Load the data

```{r}
# LOAD THE DATASET
load("data/mental_health.rda")
```

```{r data-wrangling}

# COMBINE DATASETS
red_mh <- data |>
  map_df(bind_rows)


# CLEAN THE FINAL DATASET
red_mh <- red_mh |>
  select(subreddit, author,	date, sent_neg, sent_neu, sent_pos, isolation_total, domestic_stress_total, liwc_anger, liwc_anxiety, liwc_negative_emotion, liwc_positive_emotion, liwc_sadness)


# SAVE THE FINAL DATASET
# save(red_mh, file="data/wrangled_mental_health.rda")

```

#### A (may be removed later)

#### B (may be removed later)

## Analysis



### Exploratory Data Analysis

```{r}
skim(red_mh)
```

## Data Analysis

### Question 1:

```r{}

```
-ANSWER QUESTION 1

### Question 2:

```r{}

```
-ANSWER QUESTION 2

### Question 3:

```r{}

```
-ANSWER QUESTION 3

### Question 4:

```r{}

```

-ANSWER QUESTION 4

## Modeling

-ADD MODELS HERE

### Results

-ADD RESULTS

## Conclusion

-ADD CONCLUSION
