---
title: "Replace it"
author: " Carlos van der Ley, Christine Kwon, Connor, Leonardo Gonzalez"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

## Title Trends within Mental Health Discussion on the Reddit Platform.



## Authors

Carlos van der Ley, Christine Kwon, Connor Bates , Leonardo Gonzalez

## Introduction
Mental health represents a critical aspect of an individual's overall well being as well as am important branch of Cognitive Science. Understanding not only a single individual's mental health, but also community's discussions and perceptions of mental health are essential to providing people with the help and support that they need. One online platform where community members come together to discuss mental and addiction topics is the online forum Reddit. Beginning in late 2018, the platform collected data from the platforms largest mental health centered communities for nearly two years. Our exploration of mental health discussion will leverage the rich data set provided by the platform in order to explore changes in mental health and addiction discussions in the period leading up to, as well as the beginning of the 2019-2020 pandemic.   

## Questions

**1. Following the break out of the COVID-19 Pandemic, what mental health issue saw largest spike in discussions. 

**2. How did discussions of mental health and addiction change in 2020 when compared to the years prior.  

**3. How did economic stress change over the Pandemic (Late 2019, 2020 data), and did it result in changes in a specific emotion over the other (i.e. Anger, Anxiety, General Negative Emotion)

**4. Do trends within addiction mirror mental health trends? 

## Data

### Data Introduction



### Load packages
```{r load-packages, message=FALSE}
library(tidyverse)
library(tidymodels)
library(forcats)
library(olsrr)
library(skimr)
library(srvyr)
library(osfr) #from osf website
```

### Data Import
```{r}

if(!file.exists("reddit_mental_health_report.zip")) {
  cidr <- getwd()
  mkfldr <- "data/"
  dir.create(file.path(cidr, mkfldr), recursive = TRUE)
  osf_retrieve_file("https://osf.io/7peyq/data/input/5efd0743af1156008d3b5532") |> 
    osf_download("data/") #it works
}

```

### Data Wrangling

```{r data-wrangling}

if(!file.exists("data/mental_health.zip")) {
  # GET RELATED MENTAL FILES
  mental_files = "anxiety| |socialanxiety | healthanxiety | autism | adhd | ptsd | lonely | depression | suicidewatch | bipolarreddit | bpd | addiction | alcoholism | schizophrenia"
  
  data <- list.files(recursive = TRUE,
                     path = "data/reddit_mental_health_dataset",
                     pattern = mental_files,
                     full.names = TRUE) |>
    purrr::map(~read_csv(.))
  
  # COMBINE DATASETS
  red_mh <- data |>
    map_df(bind_rows)
  
  # SELECT IMPORTANT FEATURES FROM DATASET
  red_mh <- red_mh |>
    select(subreddit, author,	date, sent_neg, sent_neu, sent_pos, isolation_total, domestic_stress_total, liwc_anger, liwc_anxiety, liwc_negative_emotion, liwc_positive_emotion, liwc_sadness)
  
  # SAVE THE FINAL DATASET
  save(red_mh, file="data/mental_health.rda")
  
  # DELETE RAW DATA: 107 FILES, +/- 2.6GB
  unlink("data/reddit_mental_health_dataset", recursive = TRUE)
}

```

#### Load the data

```{r}
# LOAD THE DATASET
load("data/mental_health.rda")
```

#### A (may be removed later)

#### B (may be removed later)

## Analysis



### Exploratory Data Analysis

```{r}
skim(red_mh)
```

## Data Analysis

### Question 1:

```r{}

```
-ANSWER QUESTION 1

### Question 2:

```r{}

```
-ANSWER QUESTION 2

### Question 3:

```r{}

```
-ANSWER QUESTION 3

### Question 4:

```r{}

```

-ANSWER QUESTION 4

## Modeling

-ADD MODELS HERE

### Results

-ADD RESULTS

## Conclusion

-ADD CONCLUSION
